---
title: "manage write xlsx in R"
author: "Haiyan Jiang"
date: "2016年12月21日"
output: html_document
---

http://stackoverflow.com/questions/21937640/handling-java-lang-outofmemoryerror-when-writing-to-excel-from-r

## R: Increase heap size for rJava
Technically speaking, this is not strictly a Unix Revelation. Still, to me it's useful enough to note this down.

We all know Java runs in a virtual machine, and assigns itself in 80% of the cases too little memory to do some real work. Result? Well, this:

java.lang.OutOfMemoryError: Java heap space
Sun gave us the -Xmx flag to increase the heap size (and -Xms to increase the stack size).

This works nicely from the commandline, the next challenge was to do something similar from an R environment. Using RWeka, which in turn depends on rJava, I found myself hitting the limit every time (which defaults to 512 MB).

The solution is to supply this -Xmx parameter before the Java Virtual Machine is initialized. From the R shell, invoke this:

```{r}
options( java.parameters = "-Xmx4g" )
# library( "RWeka" )
```
This will override the default setting and assign 4 gigabytes of heap space to the Java environment. Of course, you still need to have at least that amount of RAM.


## rexcel - issue

### What steps will reproduce the problem? 

1. run the attached script. You may have to crank up the number of sheets (ns)

### What is the expected output? What do you see instead? 
I expect an xlsx to be created, with ns tabs, each containing the same matrix of nc*nr data cells. Instead, I get an error saying "java.lang.OutOfMemoryError: Java heap space". Under other circumstances (e.g. increase the value of ns), I also get a "java.lang.OutOfMemoryError: GC overhead limit exceeded" error.

### What version of the product are you using? On what operating system?

R version 3.3.2 (2016-12-21) Platform: x86-64-unknown-linux-gnu (64-bit) xlsx-0.5.5 xlsxjars-0.5.0 rJava_0.9-6

Please provide any additional information below.

```{r}
options(java.parameters = "-Xmx1000m")
library(xlsx)
wb <- createWorkbook()

ns <- 15L
nr <- 1000L
nc <- 50L

#ns <- 10L
#nr <- 1000L
#nc <- 50L

for (i in 1:ns) {
  createSheet(wb, sheetName=paste0("Sheet", i))
}

sheets <- getSheets(wb)
x <- matrix(runif(nr*nc), nr, nc)

for (i in 1:ns) {
  sheet <- sheets[[i]]
  cb <- CellBlock(sheet, 1L, 1L, nr, nc, create=TRUE)
  CB.setMatrixData(cb, x, 1L, 1L)
}

# Error in .jcall("RJavaTools", "Ljava/lang/Object;", "newInstance", .jfindClass(class),  : 
#   java.lang.OutOfMemoryError: GC overhead limit exceeded
#   
# saveWorkbook(wb, "/Quant_Share/florent/text.xlsx")

```


options(java.parameters = "-Xmx1000m")
before library(xlsx) is called. I confirm it solved my issue. Thanks.

------------------------------------------------

his is a known issue: http://code.google.com/p/rexcel/issues/detail?id=33

While unresolved, the issue page links to a solution by Gabor Grothendieck suggesting that the heap size should be increased by setting the java.parameters option before the rJava package is loaded. (rJava is a dependency of xlsx.)

```{r}
options(java.parameters = "-Xmx1000m")
```

The value 1000 is the number of megabytes of RAM to allow for the Java heap; it can be replaced with any value you like. My experiments with this suggest that bigger values are better, and you can happily use your full RAM entitlement. For example, I got the best results using:

```{r}
options(java.parameters = "-Xmx8000m")
```
on the machine with 8GB RAM.

A further improvement can be obtained by requesting a garbage collection in each iteration of the loop. We can define a Java garbage collection function as:

```{r}
jgc <- function()
{
  gc()
  .jcall("java/lang/System", method = "gc")
}    
```




```{r}
options(java.parameters = "-Xmx2048m")  ## memory set to 2 GB
library(XLConnect)
```

allow for more memory using "options" before any java component is loaded. 
Then load XLConnect library (it uses java).
That's it. Start reading in data with readWorksheet .... and so on. :)

