---
title: "multiple ggplot"
author: "Haiyan Jiang"
date: "2016年12月20日"
output: html_document
---

To arrange multiple ggplot2 graphs on the same page, the standard R functions - par() and layout() - cannot be used.
This R tutorial will show you, step by step, how to put several ggplots on a single page.
The functions grid.arrange()[in the package gridExtra] and plot_grid()[in the package cowplot], will be used.

```{r}
# install.packages("gridExtra")
library(gridExtra)

# install.packages("cowplot")
library(cowplot)
# as follow using devtools package (devtools should be installed before using the code below):
# devtools::install_github("wilkelab/cowplot")
```

Prepare some data
```{r}
df <- ToothGrowth
# Convert the variable dose from a numeric to a factor variable
df$dose <- as.factor(df$dose)
head(df)
```

Cowplot: Publication-ready plots
The cowplot package is an extension to ggplot2 and it can be used to provide a publication-ready plots.
```{r}
library(cowplot)
# Default plot
bp <- ggplot(df, aes(x=dose, y=len, color=dose)) +
  geom_boxplot() + 
  theme(legend.position = "none")
bp
# Add gridlines
bp + background_grid(major = "xy", minor = "none")

```
Recall that, the function ggsave()[in ggplot2 package] can be used to save ggplots. However, when working with cowplot, the function save_plot() [in cowplot package] is preferred. It’s an alternative to ggsave with a better support for multi-figure plots.

```{r}
save_plot("mpg.pdf", bp, base_aspect_ratio = 1.3) # make room for figure legend

```

Arranging multiple graphs using cowplot
```{r}
# Scatter plot
head(mpg)
sp <- ggplot(mpg, aes(x=cty, y=hwy, colour=factor(cyl)) ) + 
  geom_point(size=3)
sp
# Bar plot
head(diamonds)
bp <- ggplot(diamonds, aes(x=clarity, fill = cut)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle=70, vjust=0.5))
bp

# Combine the two plots (the scatter plot and the bar plot):
plot_grid(sp, bp, labels=c("A", "B"), ncol = 2, nrow = 1)
```

The function draw_plot() can be used to place graphs at particular locations with a particular sizes. The format of the function is:

draw_plot(plot, x = 0, y = 0, width = 1, height = 1)
plot: the plot to place (ggplot2 or a gtable)
x: The x location of the lower left corner of the plot.
y: The y location of the lower left corner of the plot.
width, height: the width and the height of the plot
```{r}

```

```{r}
head(iris)
plot.iris <- ggplot(iris, aes(Sepal.Length, Sepal.Width)) + 
  geom_point() + facet_grid(. ~ Species) + stat_smooth(method = "lm") +
  background_grid(major = 'y', minor = "none") + # add thin horizontal lines 
  panel_border() # and a border around each panel
# plot.mpt and plot.diamonds were defined earlier
```



Scatter plot with marginal density plots
Step 1/3. Create some data :

```{r}
set.seed(1234)
x <- c(rnorm(500, mean = -1), rnorm(500, mean = 1.5))
y <- c(rnorm(500, mean = 1), rnorm(500, mean = 1.7))
group <- as.factor(rep(c(1,2), each=500))
df2 <- data.frame(x, y, group)
head(df2)
```

Step 2/3. Create the plots:
```{r}
# Scatter plot of x and y variables and color by groups
scatterPlot <- ggplot(df2,aes(x, y, color=group)) + 
  geom_point() +  
  scale_color_manual(values = c('#999999','#E69F00')) + 
  theme(legend.position=c(0,1), legend.justification=c(0,1))

# Marginal density plot of x (top panel)
xdensity <- ggplot(df2, aes(x, fill=group)) + 
  geom_density(alpha=.5) + 
  scale_fill_manual(values = c('#999999','#E69F00')) + 
  theme(legend.position = "none")
# Marginal density plot of y (right panel)
ydensity <- ggplot(df2, aes(y, fill=group)) + 
  geom_density(alpha=.5) + 
  scale_fill_manual(values = c('#999999','#E69F00')) + 
  theme(legend.position = "none")

```

Insert an external graphical element inside a ggplot

The function annotation_custom() [in ggplot2] can be used for adding tables, plots or other grid-based elements. The simplified format is :

annotation_custom(grob, xmin, xmax, ymin, ymax)
grob: the external graphical element to display
xmin, xmax : x location in data coordinates (horizontal location)
ymin, ymax : y location in data coordinates (vertical location)

The different steps are :

Create a scatter plot of y = f(x)
Add, for example, the box plot of the variables x and y inside the scatter plot using the function annotation_custom()
As the inset box plot overlaps with some points, a transparent background is used for the box plots.

```{r}
# Create a transparent theme object
transparent_theme <- theme(
 axis.title.x = element_blank(),
 axis.title.y = element_blank(),
 axis.text.x = element_blank(), 
 axis.text.y = element_blank(),
 axis.ticks = element_blank(),
 panel.grid = element_blank(),
 axis.line = element_blank(),
 panel.background = element_rect(fill = "transparent",colour = NA),
 plot.background = element_rect(fill = "transparent",colour = NA))

```

```{r}


set.seed(1234)
x <- c(rnorm(500, mean = -1), rnorm(500, mean = 1.5))
y <- c(rnorm(500, mean = 1), rnorm(500, mean = 1.7))
group <- as.factor(rep(c(1,2), each=500))
df2 <- data.frame(x, y, group)
head(df2)

# Scatter plot of x and y variables and color by groups
scatterPlot <- ggplot(df2,aes(x, y, color=group)) + 
  geom_point() +  
  scale_color_manual(values = c('#999999','#E69F00')) + 
  theme(legend.position=c(0,1), legend.justification=c(0,1))


p1 <- scatterPlot # see previous sections for the scatterPlot
# Box plot of the x variable
p2 <- ggplot(df2, aes(factor(1), x)) + 
  geom_boxplot(width=0.3) + coord_flip() + 
  transparent_theme
# Box plot of the y variable
p3 <- ggplot(df2, aes(factor(1), y)) + 
  geom_boxplot(width=0.3) + 
  transparent_theme
# Create the external graphical elements
# called a "grop" in Grid terminology
p2_grob = ggplotGrob(p2)
p3_grob = ggplotGrob(p3)
   
# Insert p2_grob inside the scatter plot
xmin <- min(x); xmax <- max(x)
ymin <- min(y); ymax <- max(y)
p1 + annotation_custom(grob = p2_grob, xmin = xmin, xmax = xmax, 
                       ymin = ymin-1.5, ymax = ymin+1.5)

# Insert p3_grob inside the scatter plot
p1 + annotation_custom(grob = p3_grob,
                       xmin = xmin-1.5, xmax = xmin+1.5, 
                       ymin = ymin, ymax = ymax)


```


```{r}

# Dummy plot
df <- data.frame(x = 1:10, y = 1:10)
base <- ggplot(df, aes(x, y)) +
  geom_blank() +
  theme_bw()

base + annotation_custom(
  grob = grid::roundrectGrob(),
  xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf)


# Inset plot
df2 <- data.frame(x = 1 , y = 1)
g <- ggplotGrob(ggplot(df2, aes(x, y)) +
  geom_point() +
  theme(plot.background = element_rect(colour = "black")))
base +
  annotation_custom(grob = g, xmin = 1, xmax = 10, ymin = 8, ymax = 10)

```


```{r}
x <- seq(0, 4 * pi, 0.1)
n <- length(x)
y1 <- 0.5 * runif(n) + sin(x)
y2 <- 0.5 * runif(n) + cos(x) - sin(x)
plot(x, y1, col = "blue", pch = 20)
points(x, y2, col = "red", pch = 20)

library(ggplot2)
df <- data.frame(x, y1, y2)
ggplot(df, aes(x, y = value, color = variable)) + 
    geom_point(aes(y = y1, col = "y1")) + 
    geom_point(aes(y = y2, col = "y2"))


```

