---
title: "question of matrix and order by factor"
author: "Haiyan Jiang"
date: "2016年12月27日"
output: html_document
---

## get rows or columns by names of matrix;
```{r}
rm(list = ls())
setwd("E:/hyjiang/keepInMindR")

myMatrix <- matrix(5:0, nrow=2)
rownames(myMatrix) <- c("A", "B")
colnames(myMatrix) <- c("C", "D", "E")
myMatrix

class(myMatrix)
attributes(myMatrix)

myMatrix["A", "D"] # row and column;
myMatrix["A", ]
myMatrix[, "D"]

m0 <- matrix(NA, 4, 2)
m0
rownames(m0)
colnames(m0)

```


## order by factor; 
```{r}
sex <- c("M","M","F","M","F","M","M","M","F")
x <- c("MARK","TOM","SUSAN","LARRY","EMMA","LEONARD","TIM","MATT","VIOLET")
name <- as.character(x)
y <- rnorm(9,8,1)
score <- data.frame(x, y, sex)
score

```

### order() takes multiple arguments, and it does just what you want:
```{r}
# z <- c(10:6)
# order(z)
# rank(z)
# sort(z)

score[order(sex, y, x),] # order not sort
with(score, score[order(sex, y, x),])

```



## How to sort a dataframe by column(s)?
I want to sort a data.frame by multiple columns in R. For example, with the data.frame below I would like to sort by column z (descending) then by column b (ascending):

```{r}
b = factor(c("Hi", "Med", "Hi", "Low"), levels = c("Low", "Med", "Hi"), ordered = TRUE)
x = c("A", "D", "A", "C"); y = c(8, 3, 9, 9); z = c(1, 1, 1, 2)
dd <- data.frame(b, x, y, z)
dd

```
You can use the order() function directly without resorting to add-on tools -- see this simpler answer which uses a trick right from the top of the example(order) code:

```{r}
## the default of order is ascending;
dd[with(dd, order(-z, b)), ]

```
Edit some 2+ years later: It was just asked how to do this by column index. The answer is to simply pass the desired sorting column(s) to the order() function:
rather than using the name of the column (and with() for easier/more direct access).

```{r}
dd[order(-dd[,4], dd[,1]), ]

```

You can also use the arrange() function as Hadley pointed out in the above thread:

```{r}
library(plyr)
arrange(dd,desc(z),b)
```

There are a lot of excellent answers here, but dplyr gives the only syntax that I can quickly and easily remember (and so now use very often):

```{r}
library(dplyr)
# sort mtcars by mpg, ascending... use desc(mpg) for descending
arrange(mtcars, mpg)
# sort mtcars first by mpg, then by cyl, then by wt)
arrange(mtcars , mpg, cyl, wt)
```


You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

